Hypercoercions and a Framework for Equivalence of Cast Calculi

### 2019-10-22 paper.tex
minor changes

\begin{corollary}[Correctness of \effCEK{C}]
  \label{thm:surely-monoidic-eval}
  Suppose $C$ is an instance of the Cast ADT
  and the \lazyD{} ADT, $C$ is a monoid,
  and $cast(T,l,T)=id(T)$.
  If $\judgetype{\emptyset}{e}{T}$ and $o : T$, then
  \[
    \text{\evalEqv{\effCEK{C}}{\ineffCEKD}}
  \]
\end{corollary}
\begin{proof}
  Proving that the left-hand side implies the right-hand side is
  trivial.  To prove $\judgeDeval{e}{o}$ implies
  $\judgeSeval{C}{e}{o}$, we first take one step in \effCEK{C}, and
  catch it up in \ineffCEKD\ by taking one or more steps in
  \ineffCEKD. As we just took a positive number of steps, we are
  closer to the halting state.
\end{proof}

@@

\todo[inline]{Standardize theorem statement -KC}
\begin{lemma}[Strong Bisimulation among $\mathcal{S}(\cdot)$]
	\label{thm:CEKS-bisim}
	Assume $C_1$ and $C_2$ implements Lazy D Cast ADT
	$s_{1} \in \text{\effCEK{C_1}}$,
	$s_{2} \in \text{\effCEK{C_2}}$,
	$s_{1} \approx s_{2}$,
	\begin{enumerate}
		\item If $s_1 = \sOOhalt{o}$ then $s_2 = \sOOhalt{o}$
		\item If $s_2 = \sOOhalt{o}$ then $s_1 = \sOOhalt{o}$
		\item If $\judgeSreduce{C}{s_1}{s_3}$ 
		then $\judgeDreduce{s_2}{s_4}$ and $s_3 \approx s_4$ for some $s_4$
		\item If $\judgeSreduce{C}{s_2}{s_4}$ 
		then $\judgeDreduce{s_1}{s_3}$ and $s_3 \approx s_4$ for some $s_3$
	\end{enumerate}
%	If 
%	$C_1$ and $C_2$ are instances of Lazy D Cast ADT,
%	$s_1, s_1' \in S(C_1)$ and $s_2,s_2' \in S(C_2)$,
%	$s_1 \approx s_2$,
%	$\judgeSreduce{C_1}{s_1}{s_1'}$, and
%	$\judgeSreduce{C_2}{s_2}{s_2'}$, then 
%	$s_1' \approx s_2'$
%	
	Or diagrammatically, 
	\[\begin{array}{clc}
	s_{1} & \longrightarrow_{\text{\effCEK{C}}} & s_1'\\
	\rotatebox[origin=c]{90}{$\approx$} 
	& & \rotatebox[origin=c]{90}{$\approx$} \\
	s_{2} & 
	\longrightarrow_{\text{\ineffCEKD}} & s_2' \\
	\end{array}\]
\end{lemma}
\begin{proof} By casing on $s_{1} \approx s_{2}$.
%
  \todo[inline]{The above sentence is not normal proof jargon.  Are
    you doing a proof by cases on $s_{1}
    \longrightarrow_{\text{\effCEK{C}}} s_1'$? -JS
\\---\\
By casing on the bisimulation relation. -KC}
%
  The key ideas of this proof are undoing sequencing with the property
  (2) of Lazy D Cast ADT, and handling all possibly uses of
  $cast(T,l,T)$ with property (3)-(11).
\end{proof}

@@

In this section we prove \lazyD\ hypercoercion is 
correct. First we define$applyCast(v,c)$ to make it an instance of Cast ADT, 
then prove that it is also Lazy D, and finally apply 
Theorem~\ref{thm:surely-lazyD-correct} to finish the proof.

\figref{hc-applyCast} defines$applyCast(v,c)$for \lazyD\ hypercoercion. 
Applying the identity cast for the 
dynamic type succeeds immediately. When applying a compound cast, we firstly 
apply the middle, then apply the tail. We denote by $r \; >>= \; f$ to mean 
that if $r$ is $\rOOsucc{v}$, the result is $f(v)$, otherwise the result 
is the failure.
$ applyMiddle(v,\ell,m)$ and $applyTail(v,t)$ are straightforward.
In the definition of $applyMiddle(v,\ell,m)$, we generalize 
shallow-consistency to compare middles and values ($v \smile m$) in the 
natural way.

>>>>

\begin{corollary}[Correctness of \effCEK{C}]
  \label{thm:surely-monoidic-eval}
  Suppose $C$ is an instance of the Cast ADT
  and the \lazyD{} ADT, $C$ is a monoid,
  and $cast(T,l,T)=id(T)$.
  If $\judgetype{\emptyset}{e}{T}$ and $o : T$, then
  \[
    \text{\evalEqv{\effCEK{C}}{\ineffCEKD}}
  \]
\end{corollary}
\begin{proof}
  Proving that the left-hand side implies the right-hand side is
  trivial.  To prove $\judgeDeval{e}{o}$ implies
  $\judgeSeval{C}{e}{o}$, we first take one step in \effCEK{C}, and
  catch it up in \ineffCEKD\ by taking one or more steps in
  \ineffCEKD. As we just took a positive number of steps, we are
  closer to the halting state.
  \todo[inline]{I might want to revise this proof. -KC}
\end{proof}

@@

\begin{lemma}[Strong Bisimulation among $\mathcal{S}(\cdot)$]
	\label{thm:CEKS-bisim}
	Assume $C_1$ and $C_2$ implements Lazy D Cast ADT
	$s_{1} \in \text{\effCEK{C_1}}$,
	$s_{2} \in \text{\effCEK{C_2}}$,
	$s_{1} \approx s_{2}$,
	\begin{enumerate}
		\item If $s_1 = \sOOhalt{o}$ then $s_2 = \sOOhalt{o}$
		\item If $s_2 = \sOOhalt{o}$ then $s_1 = \sOOhalt{o}$
		\item If $\judgeSreduce{C_1}{s_1}{s_3}$ 
		then $\judgeSreduce{C_2}{s_2}{s_4}$ and $s_3 \approx s_4$ for some $s_4$
		\item If $\judgeSreduce{C_2}{s_2}{s_4}$ 
		then $\judgeSreduce{C_1}{s_1}{s_3}$ and $s_3 \approx s_4$ for some $s_3$
	\end{enumerate}
%	
	Or diagrammatically, 
	\[\begin{array}{clc}
	s_{1} & \longrightarrow_{\text{\effCEK{C}}} & s_3\\
	\rotatebox[origin=c]{90}{$\approx$} 
	& & \rotatebox[origin=c]{90}{$\approx$} \\
	s_{2} & 
	\longrightarrow_{\text{\ineffCEKD}} & s_4 \\
	\end{array}\]
\end{lemma}
\begin{proof} By casing on $s_{1} \approx s_{2}$.
%
  \todo[inline]{The above sentence is not normal proof jargon.  Are
    you doing a proof by cases on $s_{1}
    \longrightarrow_{\text{\effCEK{C}}} s_1'$? -JS
\\---\\
By casing on the bisimulation relation. -KC}
%
  The key ideas of this proof are undoing sequencing with the property
  (2) of Lazy D Cast ADT, and handling all possibly uses of
  $cast(T,l,T)$ with property (3)-(11).
\end{proof}

@@

In this section we prove \lazyD\ hypercoercion is 
correct. First we define $applyCast(v,c)$ to make it an instance of Cast ADT, 
then prove that it is also Lazy D, and finally apply 
Theorem~\ref{thm:surely-lazyD-correct} to finish the proof.

\figref{hc-applyCast} defines $applyCast(v,c)$ for \lazyD\ hypercoercion. 
Applying the identity cast for the 
dynamic type succeeds immediately. When applying a compound cast, we firstly 
apply the middle, then apply the tail. We denote by $r \; >>= \; f$ to mean 
that if $r$ is $\rOOsucc{v}$, the result is $f(v)$, otherwise the result 
is the failure.
$ applyMiddle(v,\ell,m)$ and $applyTail(v,t)$ are straightforward.
In the definition of $applyMiddle(v,\ell,m)$, we generalize 
shallow-consistency to compare middles and values ($v \smile m$) in the 
natural way.
